---
title: "Pitchfork_2"
author: "Drew Whitehead"
date: "7/22/2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pitchfork Analysis: Preliminary EDA (part 2):

(continued...)

The second analysis provides a deeper dive into the distribution of the best new music awards, examining the distribution by genre and artist classes, year over year. The top panel of plot two shows the distribution of awards by genre for all artists for the period 2000 thru 20016, while the bottom three panels show the same distribution for each of the three artist classes. It’s clear that Rock genre is the dominant recipient of the best new music award overall, but the awards given by artist class are skewed. While Established and Semi-Established artists received a majority of the awards in thru the year 2010, the number of awards given after dropped off significantly, with New artists receiving a significant majority of the awards in not only Rock, but also Pop/R&B, Experimental and Rap. 


```{r cars, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
# libraries
library(ggplot2)
library(RSQLite)
library(xlsx)
library(dplyr)
library(plyr)

  # set working directory to sqlite db and connect
setwd("/Users/drewwhitehead/sqlite/")
con <- dbConnect(SQLite(), dbname="pitchfork.sqlite")

  # return table names
alltables <- dbListTables(con)

  # qurey each table name and assign to object
artists <- dbGetQuery(con, 'select * from artists')
genres <- dbGetQuery(con, 'select * from genres')
reviews <- dbGetQuery(con, 'select * from reviews')
labels <- dbGetQuery(con, 'select * from labels')
years <- dbGetQuery(con, 'select * from years')
dbDisconnect(con)
#content <- read.csv("Solo Projects/Exported CSV Files/content.csv", header = T)

  # create master dataframe from sqlite output
pitchfork_master <- merge(artists, genres)
pitchfork_master <- merge(pitchfork_master, reviews)
pitchfork_master <- merge(pitchfork_master, labels)
pitchfork_master <- merge(pitchfork_master, years)

  # set working directory back to normal
setwd("/Users/drewwhitehead/Documents/Northwestern/R-Studio/")

  # write cluster results to excel file
#write.xlsx(pitchfork_master, file = "Solo Projects/Pitchfork Master.xlsx")

# remove all duplicates from master file
pitchfork_master <- pitchfork_master[!duplicated(pitchfork_master$reviewid),]

  # create subset df of all reviews past '99 called mod_era
mod_era <- pitchfork_master[which(pitchfork_master[,'year']>1999 & pitchfork_master[,'year']<2017),]

  # remove global as it does not have enough records to be considered for average
mod_era <- mod_era[which(mod_era[,'genre']!='global'),]

  # create summary table of average score by genre ~ year
avg_rating_year <- as.data.frame(aggregate(mod_era[, "score"], list(mod_era$genre, mod_era$year), mean))

  # graph of avg score by genre YoY
p0 <- ggplot(data = avg_rating_year, aes(x = Group.2, y = x, colour = Group.1)) +       
  geom_line(aes(group = Group.1)) + geom_point() + xlab("Year") + ylab("Score") + 
  ggtitle("Avg Score by Genre YoY") + theme(plot.title = element_text(hjust = 0.5))

  # create list of artists with over 10 reviews
artist_count_df <- as.data.frame(table(mod_era$artist))
colnames(artist_count_df)[1] <- 'artist'
artist_count_df <- artist_count_df[artist_count_df$Freq >= 4,]
artist_count_df$established.rating <- NA
artist_count_df[artist_count_df$Freq >= 8, 3] <- 'Established'
artist_count_df$established.rating[is.na(artist_count_df$established.rating)] <- 'Semi-Established'
artist_count_df <- artist_count_df[-2]

  # join list of established artists
mod_era <- join(mod_era, artist_count_df, by = 'artist', type='left')
mod_era$established.rating[is.na(mod_era$established.rating)] <- 'New'

  # create separate df for each of establishment rating
mod_era_established <- mod_era[mod_era$established.rating == 'Established',]
mod_era_semi_established <- mod_era[mod_era$established.rating == 'Semi-Established',]
mod_era_new <- mod_era[mod_era$established.rating == 'New',]

  # create summary table of number of reviews by genre ~ year, for each of the establishment rating datasets
bnm_count_year_genre <- as.data.frame(aggregate(mod_era[, "best_new_music"], 
                                                             list(mod_era$year, mod_era$genre), sum))

  # create summary table of best new music bu genre ~ year, for each of the establishment rating datasets
established_bnm_count_year_genre <- as.data.frame(aggregate(mod_era_established[, "best_new_music"], 
                                                             list(mod_era_established$year, mod_era_established$genre), sum))
semi_established_bnm_count_year_genre <- as.data.frame(aggregate(mod_era_semi_established[, "best_new_music"], 
                                                             list(mod_era_semi_established$year, mod_era_semi_established$genre), sum))
new_bnm_count_year_genre <- as.data.frame(aggregate(mod_era_new[, "best_new_music"], list(mod_era_new$year, mod_era_new$genre), sum))


  # create plot of best new music by genre YoY
p3 <- ggplot(data = bnm_count_year_genre, aes(x = Group.1, y = x, colour = Group.2)) +       
  geom_line(aes(group = Group.2)) + geom_point() + xlab("Year") + ylab("Count of Reviews") + 
  ggtitle("All Artists: Count of Best New Music by Genre Rating YoY") + theme(plot.title = element_text(hjust = 0.5))

  # create plot of best new music by establishment rating YoY
p4 <- ggplot(data = established_bnm_count_year_genre, aes(x = Group.1, y = x, colour = Group.2)) +       
  geom_line(aes(group = Group.2)) + geom_point() + xlab("Year") + ylab("Count of Best New Music") + 
  ggtitle("Established: Count of Best New Music by Genre Rating YoY") + theme(plot.title = element_text(hjust = 0.5))

p5 <- ggplot(data = semi_established_bnm_count_year_genre, aes(x = Group.1, y = x, colour = Group.2)) +       
  geom_line(aes(group = Group.2)) + geom_point() + xlab("Year") + ylab("Count of Best New Music") + 
  ggtitle("Semi_Established: Count of Best New Music by Genre Rating YoY") + theme(plot.title = element_text(hjust = 0.5))

p6 <- ggplot(data = new_bnm_count_year_genre, aes(x = Group.1, y = x, colour = Group.2)) +       
  geom_line(aes(group = Group.2)) + geom_point() + xlab("Year") + ylab("Count of Best New Music") + 
  ggtitle("New: Count of Best New Music by Genre Rating YoY") + theme(plot.title = element_text(hjust = 0.5))


  # display plots 3 - 6
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
multiplot(p3, p4, cols=1)
multiplot(p5, p6, cols=1)
```
